---
title: "Problem set 1"
author: "Naomi Wambogo"
date: "10/5/2020"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    theme: united
    toc: yes
    toc_depth: 2
---

## Details

- Who did you collaborate with: no one
- Approximately how much time did you spend on this problem set: 5 hours
- What, if anything, gave you the most trouble: trying to create y~x function gave me an error that im not sure why.

### Instructions

For this problem set, you'll have three possible things to do.

- "### Question #" / "xxxxx" = Write your response to each question by replacing the "xxxx. 1-2 sentences is fine.

- "### Code #" = Write R code in the chunk corresponding to the instructions.

For this problem set, there are **15 questions** and **10 code** parts. If you get one wrong, it will be -3 points. 

Additional points will be deducted for problems with the formatting as outlined by the problem set 1 instructions (e.g., do not provide .html output, put data into data folder, etc.).

## Task 1

Answer the following questions about RMarkdown from the short lessons.

### Question 1

- What three (3) characters start and end the YAML information for an RMarkdown file?
Three dashes start and end YAML information for an Rmd file ---


### Question 2

- What does the code chunk parameter `echo = FALSE` do?
It will remove the code from the final report but will place a copy of the results into the report.

### Question 3

- How do you change the output type for an RMarkdown file?
Write a YAML header that explains what type of document to build from your R Markdown file. output:xxxxx_document
xxxxx

## Task 2: Creating baseline plot

### Code 1

Load the `tidyverse` package.

```{r load-packages, message=F, warning=F}
# make sure you have tidyverse already installed! (hint: install.packages())
#install.packages('tidyverse')
#library(tidyverse)

```

### Question 4

- What packages are loaded when you call `tidyverse`?
The packages loaded with tidyverse are ggplot2 for data visualisation,dplyr for data manipulation,tidyr for data tidying, readr for data import,purrr for functional programming,tibble, stringr for strings, and forcats for factors.

***

### Code 2 

Read in the `corrupt.csv` file and assign it to `corrupt`.

```{r load-data-1}
#load in the file 'data/corrupt.csv' and assign it to corrupt
# hint: use the read_csv function from the readr (tidyverse) package

```

***

### Code 3

Run the `glimpse()` function on the data to explore the column formats.

```{r glimpse}
# run the glimpse() function
?glimpse
glimpse(corrupt, width = NULL)
```

***

### Question 5

- How many rows (observations) and columns (variables) does the data frame include?

There are 704 observations and 5 variables.

***

However, it's not clear if there are duplicated records by year (i.e., this is panel data (record and time oriented)).

### Code 4

Run the `count()` function on `corrupt` and use `year` as the 2nd parameter. This will count how many records by each unique category in `year` (that is, each year)

```{r count}
count(corrupt,year,wt=NULL,sort = FALSE)
#corrupt%>%
  #group_by(year)%>%
    #summarise(count=n())
```

***

### Question 6

- How many different years does the dataset include?

There are four different years. 2012-2015

***

For simplicity, let's only keep 2015 records.

```{r filter-2015}
# uncomment out the lines below

 corrupt <- corrupt %>%
   filter(year == 2015) %>%
   na.omit()
```

Let's revise our existing `region` field. This will help us later on.

```{r region}
# uncomment out the lines below

 corrupt <- corrupt %>%
   mutate(region = case_when(
     region == "Middle East and North Africa" ~ "Middle East\nand North Africa",
     region == "Europe and Central Asia" ~ "Europe and\nCentral Asia",
     region == "Sub Saharan Africa" ~ "Sub-Saharan\nAfrica",
     TRUE ~ region)
     )
```

Let's now see how many countries we have for each region.

### Code 5

Using `dplyr` and piping (`%>%`), count the number of countries by region and assign it to the dataframe `region_count`. After running it, print it to the console by simply writing the name of the data frame.

```{r region-count-1, eval=F}
# hint: assign region_count through a group_by and summarise
# name your count variable `count` - if you don't you'll get an error later on
region_count <- corrupt%>%
  group_by(region)%>%
    summarize(count=n())   

# print to console by simply writing region_count
region_count
```

***

### Question 7

- How many total countries are in the "Asia Pacific" region?

There are 21 countries in the Asia Pacific region.

***

### Code 6

Create a scatterplot with the dataframe `corrupt` in which `cpi` is on the x axis, `hdi` is on the y axis, and the color of the points is `region`:

```{r species-count-2, eval=F}
# simple ggplot of geom_point()
ggplot(data = corrupt)+
  geom_point(mapping = aes(x=cpi, y=hdi, colour=region))
```

***

### Question 8

- What are three problems with this graph (or ways you could improve this graph)?

1. Some Points are overlapping each other, therefore it is not easy to distinguish apart. overplotting issue.

2. The regions are also hard to distinguish so we are better off using facet wrap method to have a plot for each region.

3. Some of the color for the plot are too similar that it maybe heard for some to make distinction

***

Now, let's modify our points.

First, let's reshape each point.

### Code 7

Within your `geom_point` function, add in the following fixed parameters: 

1. **size** to 2.5

2. **alpha** to 0.5

3. **shape** to 21

hint: since these three values are fixed, should the be inside or outside the aesthetics (`aes()`) function?

```{r graph-1,eval=F}
# modify the geom_point parameters
ggplot(data=corrupt)+
  geom_point(mapping = aes(x=cpi,y=hdi, color=region, size=2.5, alpha=0.5), shape=21)
```

***

### Question 9

- What does the `alpha` parameter do?

changes the opacity/transparency of the geom point. how dark or light the point can be.

***

The plot is too transparent. The issue is the parameter `color` encodes the color of the border, not the color of the point.

That's where we'll need the `fill` parameter.

### Code 8

Put these two parameters explicitly in the `aes()` function of the `geom_point()`:

1. `color = region`

2. `fill = region`

Also, make sure to remove any mention of color or fill in the `aes()` of your main `ggplot()` function.

```{r graph-2,eval=F}
# modify the ggplot function
ggplot(corrupt,aes(x=cpi,y=hdi,size=2.5,alpha=0.5),shape=21)+
  geom_point(mapping=aes(color=region,fill=region))
```

***

Last, let's temporarily save this graph as an object `g`. We can use the same `<-` (gets arrow) assignment operator. This will enable us to view the object or we can use it to build additional layers (Task 2).

### Code 9

Assign the ggplot from the previous part to `g` and then run `g` on its within the chunk.

```{r assign-g,eval=F}
# assign plot to g
g<-ggplot(corrupt,aes(x=cpi,y=hdi,size=2.5,alpha=0.5),shape=21)+
  geom_point(mapping=aes(color=region,fill=region))

# simply put "g" to view the plot
g
```

***

## Part 2: Re-designing

In this part, you'll add additional layers to our plot to re-design it.

This part is much more complicated, so your job will be easier: 

1. Remove the `eval=F` parameter from each chunk to run each chunk when `knit` your output.

2. Answer questions on interpreting what's going on.

For this, we'll use the same `g` object you created in the last chunk and slowly add more layers to the plot.

Before starting, we'll need two packages: `cowplot` and `colorspace`. You can install `colorspace` from CRAN (remember how to?). For `cowlplot`, you need the most recent version which is on GitHub.

Installing packages from GitHub is relatively straight-forward. But you need an additional package: `devtools`. You can then run the line below to install it.

### Code 10

Install `cowplot` and `colorspace` and call these libraries. Also, remove the `eval=F` parameter from each chunk to run each chunk when `knit` your output. (hint: you can do this for all parts via Edit > Replace and Find or CTRL + F)

```{r libraries,eval=F,warning=F,message=F}
#devtools::install_github("wilkelab/cowplot") # do you have devtools installed?
library(devtools)
library(cowplot) # for theme_minimal_hgrid()
library(colorspace) # for darken()
```

***

### Question 10

- What does the parameters `warning=F` and `message=F` do within the code chunk?

warning=FALSE prevents warnings that are generated by code from appearing in the finished line.

message=FALSE prevents messages that are generated by code from appearing in the finished line.

***

#### Import unique theme and font size

Modifying themes are very common in ggplot. There are a range of packages to change plot themes like [`ggtheme`](https://ggplot2.tidyverse.org/reference/ggtheme.html).

For this plot, we'll use a theme built within the `cowplot` package that is a minimal background with a horizontal grid.

```{r import-theme,eval=F}
g <- g +
  cowplot::theme_minimal_hgrid(12, rel_small = 1) # font size 12 pt throughout

g
```

### Question 11

- What does the `cowplot::` pre-fix for `theme_minimal_hgrid()` mean? When would it be necessary?

theme_minimal_hgrid() means minimal_horizontal grid theme, which only draws horizontal grid lines without axis lines.

***

#### Modify color scheme

Next, let's modify the color scheme. Colors can be represented by [hex colors](https://www.color-hex.com/).

Sometimes, color palettes come in as R packages (e.g., [`RColorBrewer`](https://cran.r-project.org/web/packages/RColorBrewer/index.html)). However, for this plot we'll manuall load up the colors.

```{r modify-color,eval=F, message = FALSE}
# Okabe Ito colors
region_cols <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#999999")

g <- g +
  scale_fill_manual(
    values = region_cols
  ) 

g
```

#### Darken colors by 30%

We can also darken the color scheme automatically through `colorspace`'s `darken()` function. 

```{r darken-color,eval=F}
g <- g +
  scale_color_manual(
    values = colorspace::darken(region_cols, 0.3)
  ) 

g
```

#### Add smoothing line.

Let's now overlay a basic regression, using the `geom_smooth()` function.

For this, we'll make the function a log transformation of x.

```{r smooth,eval=F}
g <- g +
  geom_smooth(
    aes(color = "y ~ log(x)", fill = "y ~ log(x)"),
    method = 'lm', formula = y~log(x), se = FALSE, fullrange = TRUE
  )

g
```

### Question 12

- What would be the difference in the plot if our smoothed function was `y~x` instead of `y~log(x)`?

Most life situations do not have a linear relationship which the log function will show but the `y~x` will show us a linear relationship between the variables.

***

#### Set x and y scales, move legend on top.

Let's now modify our scales, add scale labels, and modify the legend.

```{r set-scales,eval=F}
g <- g +
  scale_x_continuous(
    name = "Corruption Perceptions Index, 2015 (100 = least corrupt)",
    limits = c(10, 95),
    breaks = c(20, 40, 60, 80, 100),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name = "Human Development Index, 2015\n(1.0 = most developed)",
    limits = c(0.3, 1.05),
    breaks = c(0.2, 0.4, 0.6, 0.8, 1.0),
    expand = c(0, 0)
  ) +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.text = element_text(size = 9),
    legend.box.spacing = unit(0, "pt")
  ) +
  guides(
    fill = guide_legend(
      nrow = 1,
      override.aes = list(
        linetype = c(rep(0, 5), 1),
        shape = c(rep(21, 5), NA)
      )
    )
  )

g
```

#### Highlight select countries.

Last, let's add labels to highlight the countries.

We can use the `ggrepel` package that includes the `geom_text_repel()` function that makes sure not to overlap labels.

```{r highlight-busy,eval=F}
library(ggrepel)

# don't assign this to g
# if you do, then simply recreate g by running the "Run All Chunks Above" button
g1<-g +
  geom_text_repel(
    aes(label = country),
    color = "black",
    size = 9/.pt, # font size 9 pt
    point.padding = 0.1, 
    box.padding = .6,
    min.segment.length = 0,
    seed = 7654
  ) 
g1
```

Obviously, this is too busy. We have too many labels. 

Let's instead create a vector of countries we want to plot. We can the add in a new column that has the country name only if we want to plot it and nothing ("") otherwise.

```{r highlight, eval=F}
country_highlight <- c("Germany", "Norway", "United States", "Greece", "Singapore", "Rwanda", "Russia", "Venezuela", "Sudan", "Iraq", "Ghana", "Niger", "Chad", "Kuwait", "Qatar", "Myanmar", "Nepal", "Chile", "Argentina", "Japan", "China")

corrupt <- corrupt %>%
  mutate(
    label = if_else(country %in% country_highlight, country, "")
  )

# wow: %+%
# https://stackoverflow.com/questions/29336964/changing-the-dataset-of-a-ggplot-object

g <- g %+% 
  corrupt +
  geom_text_repel(
    aes(label = label),
    color = "black",
    size = 9/.pt, # font size 9 pt
    point.padding = 0.1, 
    box.padding = .6,
    min.segment.length = 0,
    seed = 7654
  ) 

g
```

### Question 13

- What do you think the `%+%` operator does (see the StackOverflow link)? Why is it necessary in this context?

it helps to avoid writing code over and over again so it anchors previous code in for a plot while adding new components. Concatenates character vectors.

***

#### Save as a pdf file

```{r save-plot,eval=F}
g + ggsave("corrupt.pdf", width = 8, height = 5)
```

You now have a pdf saved as this plot. By setting the width and height, it'll make your life so much easier if you need to reproduce this plot (very likely).

### Question 14

- From this graph, how would you interpret countries that are above the regression line versus countries that are below?

countries that are above the regression line have a high HDI meaning they are developed and have a low corruption rate. countries above the regression line also show to have a lower SSE because they are closer to the regression line compared to those below. Most countries under the regression line seem to be in  the Sub-Saharan African region.

***

### Question 15

 - What role does region have in the relationship between Corruption (Perception) and HDI?

-Sub-Saharan Africa region has the lowest human development index, whereas Europe and Central Asia seem to have the highest human development index.

- Most corrupted region is Sub-Saharan Africa, whereas prevalence of corruption is lowest in Europe and Central Asia.

-Some countries in Europe and Central Asia seem to have a high human development index but also a significant amount of corruption which raises questions as the quality of data and in what areas of the countries was the data collected.
***
install.packages('tinytex')
tinytex::install_tinytex()
library(tinytex)
tinytex::install_tinytex(force = TRUE)

